---
title: "Predicting credit default payments from client's payment history"
author: "Lianna Hovhannisyan, Arushi Ahuja, Taiwo Owoseni, Karanpreet Kaur"
date: "25/11/2021"
always_allow_html: true
output:
  github_document:
    toc: true
bibliography: default_prediction_refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(tidyverse)
```

```{r load model results}
# model <- read.csv("../results/...")
```

## Summary

In this project, we built a classification model using Logistic Regression to predict if credit account holders will make a default payment next month. The model was trained on features that hold information about the client's last 6 months bill and payment history, as well as several other characteristics such as: age, marital status, education, and gender. Overall, we are more interested in minimizing Type | error (predicting no default payment, when in reality the client made a default payment the following month), as opposed to Type || error (predicting default payment, when in reality no default payment was made by the client), we are using $f1$ as our primary scoring metric. Our model performed fairly well....

## Introduction

An account is considered the default when the client does not make the minimum payment for a consecutive 6 months period. Predicting potential credit default accounts is challenging but at the same time crucial for credit card companies. The default can happen for various reasons: the loss of a job, change in the financial market, personal difficulties, inability to work, health issues, need to use the extra cash for other bills, etc. All of the described can be considered as "out-of-control" from the customers' side. However, the default can also be intentional. An example of intentional default is when the client knows that they are no longer financially stable enough for the credit, but continue to use it until the bank intervenes [@Islam2018CreditDM]. The existence of such loopholes makes it essential for creditors to detect default accounts as soon as possible. In general, for creditors earlier they detect the potential default accounts, the lower the losses for the company [@SyedNor2019PersonalBP].

Here we ask if we can use a machine-learning algorithm to predict whether the customer will do default payment next month or not. The detection of default depends on extensive data profiling of customer's finance, including their age, payment history, marriage status, gender, education. For the creditors, it is most important to have the model that predicts the account's next month's status, especially if the client is going to make a default payment. The correct prediction will help creditors to plan their risk management and take according to actions before the situation gets out of control.

## Methods

### Data

The data set used in the project is created by Yeh, I. C., and Lien, C. H [@Yeh2009TheCO], and made publicly available for download in UCI Machine Learning Repository [@misc_default_of_credit_card_clients_350]. The data can be found [here](<https://archive-beta.ics.uci.edu/ml/datasets/default+of+credit+card+clients>), specifically [this file](<https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls>). The dataset is based on Taiwan's credit card client default cases from April to September. It has 30000 examples, and each example represents particular client's information. The dataset has 24 observations with respective values such as gender, age, marital status, last 6 months bills, last 6 months payments, etc, including the final default payment of next month column: labeled 1 (client will make a default) and 0 (client will not make a default).   

### Analysis

The Logistic Regression algorithm was used to build a classification model for the default credit card dataset to predict whether the client will make a default payment or not in the next month (can be found in the "default payment next month" column with binary label). All the features included in the original dataset were used for tuning and fitting the model. The hyperparameter $C$ was chosen using 10 cross-validations with $f1$ score as scoring metrics. The Python programming language [@Python] and the following Python packages were used for the model analysis: docopt [@docoptpython], sklearn [@sklearn_api], altair [@vanderplas2018altair], pandas [@mckinney2010data], [@020NumPy-Array], os [@Python], requests [@chandra2015python], pickle [@van1995python], matplot [@hunter2007matplotlib], seaborn [@michael_waskom_2017_883859]; as well as R programming language [@R] for generating this report and the following packages: knitr [@knitr], tidyverse [@tidyverse]. The code used to perform the analysis and create the report can be found [here](https://github.com/UBC-MDS/credit_default_prediction).

### Results & Discussion

-   EDA

- Analysis

## References
