---
title: "Predicting credit default payments from client's payment history"
author: "Lianna Hovhannisyan, Arushi Ahuja, Taiwo Owoseni, Karanpreet Kaur"
date: "25/11/2021"
always_allow_html: true
output:
  html_document:
    toc: true
bibliography: default_prediction_refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(tidyverse)
```

```{r load model results}
# model <- read.csv("../results/...")
```

## Summary




## Introduction

An account is considered default, when the client does not make minimum payment for consecutive 6 months period. Predicting potential credit default accounts is challenging, but at the same time crucial for credit card companies. The default can happen for various reasons: the loss of job, change in financial market, personal difficulties, inability to work, health issues, need to use the extra cash for other bills, etc. All of the described can be considered as "out-of-control" from the customers side. However, the default can also be intentional. An example of intentional default is when the client knows that they are no longer financially stable enough for the credit, but continue to use it until the bank intervenes [@Islam2018CreditDM]. The existence of such loopholes makes it essential for creditors to detect default accounts as soon as possible. In general, for creditors earlier they detect the potential default accounts, the lower the losses for the company [@SyedNor2019PersonalBP].

Here we ask if we can use a machine learning algorithm to predict whether customer will do default payment next month or not. The detection of default depends on extensive data profiling of customer's finance, including their age, payment history, marriage status, gender, education. For the creditors, it is at most importance to have the model that predicts the account's next month status, especially if the client is going to make a default payment. The correct prediction will help creditors to plan their risk management and take according actions before the situation gets out of control.

## Methods

### Data

The data set used in the project is created by Yeh, I. C., and Lien, C. H [@Yeh2009TheCO], and made publicly available for download in UCI Machine Learning Repository [@misc_default_of_credit_card_clients_350]. The data can be found [here](<https://archive-beta.ics.uci.edu/ml/datasets/default+of+credit+card+clients>), specifically [this file](<https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls>). The dataset is based on Taiwan's credit card client default cases from April to September. It has 30000 examples, and each example represents one client information. The dataset has 24 observations with respective values such as gender, age, marigal status, last 6 months bills, last 6 months payments, etc, including the final default payment of next month column: labeled 1 (client will make a default) and 0 (client will not make a default).   

### Analysis

The Logisitic Regression algorithm was used to build a classification model for the default credit card dataset to predict whether the client will make a default payment or not in the next month (can be found in "default payment next month" column with binary label). All the features included in the original dataset were used for tuning and fitting the model. The hyperparameter $C$ was chosen using 10 cross validation with f1 score as scoring metrics. The Python programming language [@Python] and the following Python packages were used for the model analysis: docopt [@docoptpython], sklearn [@sklearn_api], altair [@vanderplas2018altair]; as well as R programming language [@R] for generating this report and the following packages: knitr [@knitr], tidyverse [@tidyverse]. 

### Results & Discussion

-   EDA

- Analysis

## References
